\documentclass{letter}

\signature{Alex Groce}
%\address{21 Bridge Street \\ Smallville \\ Dunwich DU3 4WE}
\begin{document}

\begin{letter}{}
\opening{Dear reviewers:}

I submit this revision to my Onward! Essay, which I believe addresses the concerns noted by reviewers.

The proposed changes from the meta-review are:

\begin{enumerate}

\item{\bf Improve the introduction to explain early (Section 1.0) why the essay continues at all:}

\item{\bf Better articulate the relationship between this short description of MDD and the "bigger picture," and the relationship between Sections 1 and 5}

\item{\bf Double check places humor may have not translated well to writing (examples in Review B and below).}

\end{enumerate}

The latexdiff pdf that follows shows how these concerns were addressed.  I think the final paragraph added to the introduction handles the first concern.  The second concern (the most difficult) I have addressed in two ways.  Narrowly, I added a paragraph near the end of section 4 that explains why I think \emph{in small MDD-appropriate scopes} this will often not be a challenge:  namely, surviving mutants under the level of coverage and oracle power TDD creates even without mutation will tend to cluster around similar causes for being missed.  This doesn't address equivalent mutants, but the traditional 4-10\% of mutants that are equivalent probably aren't a huge burden at a scope suitable for TDD/MDD.  Second, I think the new section about ``the River'' of software development gives better insight into why I think MDD might work, and how it might work, that connects better to Beck's pyschological emphasis.

Finally, I removed weak humor and made it clear where a gentle joke was being made about what an essay ``is.''

\closing{Sincerely,}



\end{letter}
\end{document}